Analysis of Crime Statistics from 2014 to 2023 in Neighborhoods of Varying Home Price

### 2024-09-23 paper/paper.qmd
Removed redundant files, updated README, and LLM usage file.

**Data Selection Reasoning**: While Toronto’s census data also contains valuable socio-economic information, the housing price dataset was chosen because it was updated more frequently and provides clearer spatial links between neighborhoods and socio-economic status. The census data, recorded in 2015 and 2021, presented two key limitations for this analysis:

@@

- Autotheft displays an exponential trend, increasing drastically from 2015 levels.

@@

- Shooting has increased by over 300% since 2015 levels

@@

- Theft from motor vehicles rates has remained around 300 per 100,000 from 2015 to 2023.

- Robbery rates have decreased significantly from 2015 to 2023 by about 34%.

@@

- Break and enter rates have remained relatively stagnant from 2015, with dips corresponding to Covid-19 lockdowns.

@@

```
Figure 4 shows that the overall crime rate has increased at a much higher rate in the top 10 highest price neighbourhoods than in the bottom 10 lowest price neighbourhoods. Both the highest and lowest priced neighbourhoods have experienced a steady increase in overall crime rates from 2015 to 2023, with the highest crime rate being that of the top 10 highest price neighbourhoods in 2023.

@@

- **Autotheft**: Bottom 10 Lowest neighborhoods show a steady increase in autotheft rates from 2015 onward, peaking around 2022. Top 10 Highest neighborhoods have significantly lower autotheft rates throughout the period, but they also show a slight upward trend, though the increase is far more gradual.

- **Biketheft**: In the earlier years (2015–2017), Bottom 10 Lowest neighborhoods have higher bike theft rates, but Top 10 Highest neighborhoods catch up around 2018–2020. From 2020 onwards, the bike theft rates in Top 10 Highest neighborhoods are actually higher than those in the Bottom 10 Lowest neighborhoods.

>>>>

**Data Selection Reasoning**: While Toronto’s census data also contains valuable socio-economic information, the housing price dataset was chosen because it was updated more frequently and provides clearer spatial links between neighborhoods and socio-economic status. The census data, recorded in 2014 and 2021, presented two key limitations for this analysis:

@@

- Autotheft displays an exponential trend, increasing drastically from 2014 levels.

@@

- Shooting has increased by over 300% since 2014 levels

@@

- Theft from motor vehicles rates has remained around 300 per 100,000 from 2014 to 2023.

- Robbery rates have decreased significantly from 2014 to 2023 by about 34%.

@@

- Break and enter rates have remained relatively stagnant from 2014, with dips corresponding to Covid-19 lockdowns.

@@

```
Figure 4 shows that the overall crime rate has increased at a much higher rate in the top 10 highest price neighbourhoods than in the bottom 10 lowest price neighbourhoods. Both the highest and lowest priced neighbourhoods have experienced a steady increase in overall crime rates from 2014 to 2023, with the highest crime rate being that of the top 10 highest price neighbourhoods in 2023.

@@

- **Autotheft**: Bottom 10 Lowest neighborhoods show a steady increase in autotheft rates from 2014 onward, peaking around 2022. Top 10 Highest neighborhoods have significantly lower autotheft rates throughout the period, but they also show a slight upward trend, though the increase is far more gradual.

- **Biketheft**: In the earlier years (2014–2017), Bottom 10 Lowest neighborhoods have higher bike theft rates, but Top 10 Highest neighborhoods catch up around 2018–2020. From 2020 onwards, the bike theft rates in Top 10 Highest neighborhoods are actually higher than those in the Bottom 10 Lowest neighborhoods.
